---
title: "Alaska 2020-2023 NOAA OME PMEL Sample Map"
output: html_document
date: "2023-04-18"
---
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r, echo=FALSE, hide=TRUE, warnings=FALSE,message=FALSE}
library(tidyverse)
library(here)
library(sf)
library(devtools)
#devtools::install_github("katiejolly/nationalparkcolors")
library(nationalparkcolors)
library(knitr)
library("rnaturalearth")
library("rnaturalearthdata")
library("readxl")
library(parzer)

world <- ne_countries(scale = "medium", returnclass = "sf")

here()
```

```{r,echo=FALSE, hide=TRUE, warnings=FALSE,message = FALSE}
sample_data <- read_excel(here("data","Alaska_Arctic_eDNASampleLocations.xlsx"))
ome_data <- read_excel(here("data","20240202_OME_Sample_Spreadsheet.xlsx"))

sample_data <- sample_data %>% 
  rename(., Sample_ID="Sample NAME",
         Cruise ="Cruise/Expedition Name",
         Site="Site Name",
         Depth_m="Depth (m)") %>% 
  mutate(., Lat_deg=parse_lat(Latitude),
         Lon_deg=parse_lon(Longitude)) %>% 
  mutate(`Depth_m`=as.numeric(`Depth_m`))

ome_data <- ome_data %>% 
  rename(., Sample_ID="FINAL Sample NAME") %>% 
  dplyr::select(Sample_ID,Date_utc=`Collection Date (UTC)`,Date_local=`Collection Date (local)`)


attr(ome_data$Date_utc, "tzone") <- "UTC"
attr(ome_data$Date_local, "tzone") <- "US/Alaska"

ome_data %>% 
  mutate(., Date = if_else(is.na(Date_utc), Date_local, Date_utc)) %>% 
  dplyr::select(Sample_ID,Date)-> dates_clean

sample_data %>% 
  left_join(dates_clean) %>% 
  mutate(., Year = as.numeric(format(Date, format="%Y"))) %>% 
  mutate(., Date_easy = format(Date, format="%m-%d-%Y")) %>% 
  mutate(., Month = as.numeric(str_sub(Date_easy,1L,2L)),
         Season = case_when(Month<4~"Winter",
                            Month <7~"Spring",
                            Month <9~"Summer",
                            Month < 12~"Fall")) %>% 
  mutate(., Year = case_when(Sample_ID == "E26.1B.DY2012"~ 2020,
       Sample_ID == "E1953.SKQ23-12S"~ 2023, 
      Sample_ID == "E1954.SKQ23-12S"~ 2023,
      Sample_ID == "E1955.SKQ23-12S"~ 2023,
      TRUE ~Year)) %>% 
  mutate(., Season = case_when(Sample_ID == "E26.1B.DY2012"~ "Fall",
       Sample_ID == "E1953.SKQ23-12S"~ "Fall", 
      Sample_ID == "E1954.SKQ23-12S"~ "Fall",
      Sample_ID == "E1955.SKQ23-12S"~ "Fall",
      TRUE ~Season))-> alaska_samples


```
# General Statistics
```{r, warnings=FALSE,echo=FALSE, hide=TRUE, warnings=FALSE,message = FALSE}
#make some colors


pal <- park_palette("Denali", 5)
pal2 <- park_palette("Yellowstone", 5)

pal3 <- park_palette("Arches", 5)
pal4 <- park_palette("Hawaii", 5)
```



### Unique Samples, Cruises, and Sites:
```{r,echo=FALSE}

alaska_samples %>% 
dplyr::summarise(n_distinct(`Sample_ID`), n_distinct(Cruise), n_distinct(`Site`))

```

## Depth Distribution of Samples:

```{r,echo=FALSE,warning=FALSE}
my_binwidth <- 1

alaska_samples %>%  
  group_by(`Depth_m`) %>% 
  ggplot(., aes(x=`Depth_m`))+
  geom_histogram(binwidth = my_binwidth, colour= "black", fill = pal[1]) +
  geom_density(aes(y = ..density.. * (nrow(sample_data) * my_binwidth)), fill=pal[1], alpha = .4) +theme_bw() + scale_x_sqrt(breaks=c(0,10,50,100,200,400,600)) +xlab("Depth (m)") + ylab("Count")
```
  
Vast majority of Alaska samples are taken within the top 100m.

# Cruise Map

```{r,echo=FALSE,warning=FALSE, fig.height= 8, fig.width= 12}

min_lat <- min(alaska_samples$Lat_deg)
max_lat <- max(alaska_samples$Lat_deg)

min_lon <- min(alaska_samples$Lon_deg)
max_lon <- max(alaska_samples$Lon_deg)

alaska_samples$Year <- as.character(alaska_samples$Year)


ggplot(data = world) +
    geom_sf() +
    geom_point(data = alaska_samples, aes(x = Lon_deg, y = Lat_deg, colour=Year)) +
    coord_sf(xlim = c(min_lon-2, max_lon+2), ylim = c(min_lat-2, max_lat+2), expand = FALSE) +theme_bw() +xlab("Longitude") +ylab("Latitude") +scale_color_manual(values=c(pal[1],pal2[1],pal2[5],pal3[1])) +
  geom_text(data = alaska_samples, aes(x = Lon_deg, y = Lat_deg, label=Site),size = 0.5) +facet_grid(Season~Year)


```

Sikuliaq stations from 2021 were not labeled with site in the .xlsx file so the names are missing.

Fall sites have far more repeated coverage. 2022 had lowest coverage.

# Cruise Map Size Proportional to Years Surveyed
```{r,echo=FALSE,warning=FALSE, fig.height= 8, fig.width= 12}

alaska_samples %>% 
  select(Lon_deg,Lat_deg, Year, Season) %>% 
  mutate(., Lon_deg_r=round(Lon_deg,1),
          Lat_deg_r=round(Lat_deg,1)) %>% 
    select(Lon_deg_r,Lat_deg_r, Year, Season) %>% 
  distinct() %>% 
  group_by(Lon_deg_r,Lat_deg_r, Season) %>% 
  count() -> group_by_site

ggplot(data = world) +
    geom_sf() +
    geom_point(data = group_by_site, aes(x = Lon_deg_r, y = Lat_deg_r, size=n),alpha=0.6, colour=pal[1]) +
    coord_sf(xlim = c(min_lon-2, max_lon+2), ylim = c(min_lat-2, max_lat+2), expand = FALSE) +theme_bw() +xlab("Longitude") +ylab("Latitude") +facet_grid(.~Season)


```

This is a count of the stations (had to round to nearest .1 degree lattitude and longitude since missing site/station names). Only 1 spring site was visited in 2022 and 2023 There are a handful of sites visited 3 or 4 times in the Fall since 2020. 

## Heatmap of Sampling Intensity
```{r,echo=FALSE,warning=FALSE, fig.height= 8, fig.width= 12}

alaska_samples %>% 
  select(Lon_deg,Lat_deg, Year, Season, Depth_m) %>% 
  mutate(., Lon_deg_r=round(Lon_deg,1),
          Lat_deg_r=round(Lat_deg,1)) %>% 
    select(Lon_deg_r,Lat_deg_r, Year, Season, Depth_m) %>% 
  distinct() %>% 
  group_by(Lon_deg_r,Lat_deg_r, Year, Season) %>% 
  count() %>% 
  mutate(., `Distinct Depths Sampled`=n) %>% 
  mutate(., Lat_Lon = str_c(Lat_deg_r, "_",Lon_deg_r)) %>% 
  ggplot(., aes(x=Year, y= Lat_Lon, color=`Distinct Depths Sampled`, fill=`Distinct Depths Sampled`)) +geom_tile() +facet_grid(.~Season)


```
### Distribution of Sites Revisited in Fall
```{r,echo=FALSE,warning=FALSE}

alaska_samples %>% 
  select(Lon_deg,Lat_deg, Year, Season) %>% 
  mutate(., Lon_deg_r=round(Lon_deg,1),
          Lat_deg_r=round(Lat_deg,1),
         ) %>% 
    select(Lon_deg_r,Lat_deg_r, Year, Season) %>% 
  distinct() %>% 
  group_by(Lon_deg_r,Lat_deg_r, Season) %>% 
  count() %>% 
    mutate(., `Distinct Years Sampled`=n) -> distinct_years_sampled



distinct_years_sampled %>% 
  filter(., Season =="Fall") %>% 
   ggplot(., aes(x=`n`))+
  geom_histogram(binwidth = my_binwidth, colour= "black", fill = pal[4]) +
  geom_density(aes(y = ..density.. * (nrow(sample_data) * my_binwidth)*.15), fill=pal[4], alpha = .4) +theme_bw()  + ylab("Count")
  
```

### Distribution of Sites Revisited in Spring

```{r,echo=FALSE,warning=FALSE}

distinct_years_sampled %>% 
  filter(., Season =="Spring") %>% 
   ggplot(., aes(x=`n`))+
  geom_histogram(binwidth = my_binwidth, colour= "black", fill = pal[1]) +
  geom_density(aes(y = ..density.. * (nrow(sample_data) * my_binwidth*.03)), fill=pal[1], alpha = .4) +theme_bw()  + ylab("Count")
```




